
---
title: "Assignment-1"
author: "Md Kamrul Hasan Khan"
date: "January 27, 2020"
output: pdf_document
---

**Homework must be submitted as a single pdf file which is the output of the RMarkdown file. Please save the file as `Last-Name-assignment-1.pdf` and email it to** jrtipton@uark.edu. For example, the student with the last name of Fox would submit the file `Fox-assignment-1.pdf`.

The objective of this assignment is to familiarize yourself with data. In the assignment, you will solve problems using `RMarkdown`. 

# Summary of Data

The data provided in _USHCN.RData_ posted on the website consists of multiple objects from the United States Historical Climatology Network
([http://cdiac.ornl.gov/epubs/ndp/ushcn/ushcn.html](http://cdiac.ornl.gov/epubs/ndp/ushcn/ushcn.html)).

```{r}
load(here::here("data", "USHCN.RData"))
```


There are $n$ `r nrow(precip)` observations of daily precipitation measurements (in hundredths of inches) at $n_s$ `r ncol(precip)` sites in the $n \times n_s$ matrix _precip_. The $n_s \times 2$ matrix _lon_lat_ contains the longitude (first column) and latitude (second column) for each of the $n_s$ stations. The $n$-vector _days_ contains the day of the year (1-365) of the measurement (ignoring leap-days) and the $n$-vector _years_ contains the year of the measurement. The $n_s$ vector _elev_ contains the elevation of each station and the $n_s$ vector _station_name_ contains the name of each station. 

# Problem 1

Create a single data.frame that is in long format (each row is an observation, each column is a variable -- the data.frame should have `r nrow(precip) * ncol(precip)` rows) that merges the objects 
_precip_, _lon_lat_, _days_, _years_, _elev_, and _station_name_. For examples of _tidy_ data , see [https://tidyr.tidyverse.org/articles/tidy-data.html](https://tidyr.tidyverse.org/articles/tidy-data.html) and [https://tidyr.tidyverse.org/](https://tidyr.tidyverse.org/)

# Solution of Problem 1

A single data.frame creation:

```{r}
library(dplyr)
library(tidyverse)
library(tidyr)

load(here::here("data", "USHCN.RData"))

d1 <- data.frame(precip, days, years)
n = nrow(precip)
d2 <- d1 %>% 
	gather(precip, precipitation, X1:X1218) %>%
	select(-precip) %>%
	mutate(longitue = rep(lon_lat[ , 1], each = n) ) %>%
	mutate(latitude = rep(lon_lat[ , 2], each = n) ) %>%
	mutate(elevation = rep(elev, each = n) ) %>%
	mutate(station = rep(station_name, each = n) )

```


# Problem 2

For each station, calculate the total annual rainfall at each site. Plot these annual summaries as a ["spaghetti" plot](https://stackoverflow.com/questions/17490515/spaghetti-plot-with-different-colors-for-each-subject-in-ggplot2) where each station is given its own line. 

# Solution of Problem 2

```{r}

d3 <- d2 %>%
   	group_by(years, station) %>%
   	summarize(total_precip = sum(precipitation,  na.rm = TRUE))

d3 %>% 
	ggplot(aes(x = years, y = total_precip, col = station)) +
	geom_line() +
	xlab("Years")+
	ylab("Total Precipitation") +
	labs(title = "Stationwise Year vs Total Precipitation Plot")+
	theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

```




# Problem 3

Choose 3 sites from the data. Calculate a linear regression of daily precipitation vs. year at each of the sites. Provide figures and write a few sentences describing the results. Make sure you write out the answers in **clear, English sentences. Communication matters!**


# Solution of Problem 3

This data contains significant proportion (around $17\%$) of missing data. Hence below I choose three sites with least missing data ($0\%, 0.02\%, 0.06\%$). Then are linear regression model is run for each station independently where the response constructed by taking average of all year's precipitations at each day and the covariate is those yearly precipitations on the corresponding day. Here the reponse is actully a function of covriates and for that reason all the regression coefficents are simmilar. The residual plots in next three figures shows that the residuals are significantly small of all stations. This is because the response can be written as a function of all covariates. The qqplots shows that the data follows approximately normal distribution.

```{r}

aa = apply(precip, 2, function(x) sum(x >=0, na.rm = TRUE))
bb = sort(aa, decreasing = TRUE)

st_1 <- station_name[which(aa == bb[1])]
st_2 <- station_name[which(aa == bb[2])]
st_3 <- station_name[which(aa == bb[3])]

d4 <- d2 %>%
  group_by(days, station) %>%
  summarize(total_precip = mean(precipitation,  na.rm = TRUE))

## For first station 

yy_1 <- d4 %>%
  filter(station == st_1)%>%
  pull(total_precip)

d_st_1 <- d2 %>%
  pivot_wider(names_from = years, values_from = precipitation) %>%
  filter(station == st_1) %>%
  select(-(1:5)) %>%
  mutate(response = yy_1)

station_1_lm = lm( response ~. , data = d_st_1 )
par(mfrow = c(2, 1))
plot(station_1_lm, which = 1:2)

## For second station 

yy_2 <- d4 %>%
  filter(station == st_2)%>%
  pull(total_precip)

d_st_2 <- d2 %>%
  pivot_wider(names_from = years, values_from = precipitation) %>%
  filter(station == st_2) %>%
  select(-(1:5)) %>%
  mutate(response = yy_2)

station_2_lm = lm( response ~. , data = d_st_2 )
par(mfrow = c(2, 1))
plot(station_2_lm, which = 1:2)


## For third station 

yy_3 <- d4 %>%
  filter(station == st_3)%>%
  pull(total_precip)

d_st_3 <- d2 %>%
  pivot_wider(names_from = years, values_from = precipitation) %>%
  filter(station == st_3) %>%
  select(-(1:5)) %>%
  mutate(response = yy_3)

station_3_lm = lm( response ~. , data = d_st_3 )
par(mfrow = c(2, 1))
plot(station_3_lm, which = 1:2)



```


# Problem 4

Create a figure that displays the data. This can be a map of raw data, summarized data (i.e. like in problem 2), animation, time series plot, etc. that you think describes an interesting feature in the data. Write a few sentences that explain your plot and what makes it interesting. Make sure you write out the answers in **clear, English sentences. Communication matters!**


# Solution of Problem 4

Since this data has a significant proportion of NAs, a summary data representation may be more informative than representing the whole raw data. Moreover, the temporal patern may explain the data more clearly. Therefore,
the next figure dipicts the plot of station wise daily precipitation averaged over the years to show the patern of the precipitation within years. The figure represents that although in most of the station the averaged precipitation remains similar over the year, in some station there is significant jumps at the middle of the year.

```{r}
d2 %>%
  group_by(days, station) %>%
  summarize(mean_precip = mean(precipitation,  na.rm = TRUE))%>%
	ggplot(aes(x = days, y = mean_precip, col = station)) +
	geom_line() +
	ylab("Average Precipitation") +
	xlab("Days") +
	labs(title = "Stationwise Days vs Average Precipitation Plot")+
	theme(legend.position = "none", plot.title = element_text(hjust = 0.5))

```




