---
title: "Assignment 2"
author: "Md Kamrul Hasan Khan"
date: "Due 5PM on Friday February 14, 2020"
output: pdf_document
urlcolor: blue
---

```{r setup, include=FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
Make a map of the total annual precipitation (The sum of annual precipitation from columns 6-17).

\textbf{Answer:}

```{r, message = FALSE,  warning = FALSE}
if (!require("devtools")) install.packages('devtools')
#if (!require("rspatial")) devtools::install_github('rspatial/rspatial')
library(tidyverse)
library(fields)
```

```{r}
dat <- readRDS(here::here("data","precipitation.Rds"))
dat <- dat %>%
            mutate(annual_total = rowSums(dat[ , 6:17])) %>%
            mutate(log_annual_total = log(annual_total))
glimpse(dat)
```

```{r}
dat %>%
    ggplot(aes(x = 1:length(annual_total), y = log(annual_total) )) +
    geom_line()+
    geom_point() +
    xlab("")+
    ylab("Log Total Annual Precipitation")+
    ggtitle("Figure 1: Log Total Annual Precipitation") +
    theme(plot.title = element_text(hjust = 0.5))
 
quilt.plot(dat$LONG, dat$LAT, dat$log_annual_total, nx = 30, ny = 30, 
           xlab = "Longitude", ylab = "Latitude", 
           main = "Figure 2: Log Total Annual Precipitation") 
points(dat$LONG, dat$LAT, cex = .3)
```

Figure 1 shows that trace plot and Figure 2 depicts the spatial representation of log total annual precipitation in California. Figure 2 shows that the north western part has more precipitation compared to the south eastern part.

# Problem 2

Using the total annual rainfall, fit 8 different spatial models using the _geoR_ package. 4 models with the spatial covariates (lat, long, and altitude) and 4 models without the covariates (mean is intercept-only). For each of the 4 models, fit two models with a nugget and two models without a nugget. Finally, for each model, consider a Matern covariance and an exponential covariance. Make sure you report the AIC and BIC for each of the models in a table. The 8 models are given in the table below. 

```{r}
knitr::kable(
    data.frame(
        Model      = 1:8, 
        Covariates = rep(c("Spatial Coefficients", "Intercept Only"), each = 4),
        Covariance = rep(rep(c("Matern", "Exponential"), each = 2), times = 2),
        Nugget     = rep(rep(c("No Nugget", "Nugget"), times = 2), times = 2)
        
    )
)
```

\textbf{Answer:}

Precipitation cannot take any negative values. Hence, total annual precipitation is modeled in log scale. Before fitting the data using any spatial model, I first fit the log total annual precipitation using a multiple linear regression where the predictors are latitude, longitude and altitude and check the distribution of the residuals. Figure 3 shows the histogram of the residuals. The histogram depicts that the residuals follow approximately normal distribution.

```{r}

dat %>%
    mutate(residual = resid(lm(log_annual_total ~ LAT + LONG + ALT, data = .))) %>%
    ggplot(aes(residual)) +
    geom_histogram(bins = 100) +
    xlab("Residual") +
    ylab("Frequency") +
    ggtitle("Figure 3: Histogram of Residuals")+
    theme(plot.title = element_text(hjust = 0.5))

```

```{r, message = FALSE,  warning = FALSE}
#if (!require("groR")) install.packages("geoR")

library(geoR)

fit_model_1 <- likfit(
    data         = dat$log_annual_total,
    trend        = ~ dat$LAT + dat$LONG + dat$ALT,
    coords       = cbind(dat$LAT, dat$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    fix.nugget = TRUE, nugget = 0,
    message = FALSE
)
#summary(fit_model_1)

fit_model_2 <- likfit(
    data         = dat$log_annual_total,
    trend        = ~ dat$LAT + dat$LONG + dat$ALT,
    coords       = cbind(dat$LAT, dat$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    message = FALSE
)
#summary(fit_model_2)

fit_model_3 <- likfit(
    data         = dat$log_annual_total,
    trend        = ~ dat$LAT + dat$LONG + dat$ALT,
    coords       = cbind(dat$LAT, dat$LONG),
    cov.model    = "exponential",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    fix.nugget = TRUE, nugget = 0,
    message = FALSE
)
#summary(fit_model_3)

fit_model_4 <- likfit(
    data         = dat$log_annual_total,
    trend        = ~ dat$LAT + dat$LONG + dat$ALT,
    coords       = cbind(dat$LAT, dat$LONG),
    cov.model    = "exponential",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    message = FALSE
)
#summary(fit_model_4)

fit_model_5 <- likfit(
    data         = dat$log_annual_total,
    coords       = cbind(dat$LAT, dat$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    fix.nugget = TRUE, nugget = 0,
    message = FALSE
)
#summary(fit_model_5)

fit_model_6 <- likfit(
    data         = dat$log_annual_total,
    coords       = cbind(dat$LAT, dat$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    message = FALSE
)
#summary(fit_model_6)

fit_model_7 <- likfit(
    data         = dat$log_annual_total,
    coords       = cbind(dat$LAT, dat$LONG),
    cov.model    = "exponential",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    fix.nugget = TRUE, nugget = 0,
    message = FALSE
)
#summary(fit_model_7)

fit_model_8 <- likfit(
    data         = dat$log_annual_total,
    coords       = cbind(dat$LAT, dat$LONG),
    cov.model    = "exponential",
    ini.cov.pars = c(var(dat$log_annual_total), 1),
    message = FALSE
)
#summary(fit_model_8)
```

```{r}
AIC <- rep(0, times = 8)
BIC <- rep(0, times = 8)
log_likelihood <- rep(0, times = 8)

for(i in 1:8) 
    {
    log_likelihood[i] <- round(eval(parse(text=paste0("fit_model_", i, "$loglik"))) , 3)
    AIC[i] <- round(eval(parse(text=paste0("fit_model_", i, "$AIC"))) , 3)
    BIC[i] <- round(eval(parse(text=paste0("fit_model_", i, "$BIC"))) , 3)
    }

knitr::kable(
    data.frame(
        Model      = 1:8, 
        Covariates = rep(c("Spatial Coefficients", "Intercept Only"), each = 4),
        Covariance = rep(rep(c("Matern", "Exponential"), each = 2), times = 2),
        Nugget     = rep(rep(c("No Nugget", "Nugget"), times = 2), times = 2),
        Log_likelihood = log_likelihood,
        AIC = AIC,
        BIC = BIC
    ),
    caption = "Model comparison using information based criterion"
)

```

There are 8 candidate models to fit the log total annual precipitation. After fitting the data the models are compared using log likelihood as well as several information based criteria like AIC and BIC. Model comparison outputs are shown in Table 2. All three criteria show that adding spatial covariates (latitude, longitude and altitude) in the model improves the outputs significantly and Model 2 (model with spatial coefficients, Matern covariance and nugget) performs significantly better compared to the others because the log likelihood is maximum, and AIC as well as BIC are minimum.

# Problem 3

Plot the fitted correlation function for each of the 8 models. Is there evidence of residual spatial autocorrelation? Does the inclusion of the covariates (lat, long, and altitude) have an effect on the correlation structure? **Write a few sentences to interpret the results.** Make sure you write out the answers in **clear, English sentences. Communication matters!**

\textbf{Answer:}

```{r, message = FALSE,  warning = FALSE}
beta_0 <- rep(0, times = 8)
partial_sill <- rep(0, times = 8)
phi <- rep(0, times = 8)


for(i in 1:8) {
    beta_0[i] <- round(eval(parse(text=paste0("fit_model_", i, "$beta")))[1] , 3)
    partial_sill[i] <- round(eval(parse(text=paste0("fit_model_", i, "$sigmasq"))), 3)
    phi[i] <- round(eval(parse(text=paste0("fit_model_", i, "$phi"))) , 3)
    }

beta_1 <- rep(0, times = 8)

beta_2 <- rep(0, times = 8)

beta_3 <- rep(0, times = 8)

for(i in 1:4) 
    {
    beta_1[i] <- round(eval(parse(text=paste0("fit_model_", i, "$beta")))[2] , 3) 
    beta_2[i] <- round(eval(parse(text=paste0("fit_model_", i, "$beta")))[3] , 3)
    beta_3[i] <- round(eval(parse(text=paste0("fit_model_", i, "$beta")))[4] , 4)
    }

nugget <- rep(0, times = 8)

for(i in c(2,4,6,8)) {
    nugget[i] <- round(eval(parse(text=paste0("fit_model_", i, "$nugget"))) , 4)}

kappa <- rep(0.5, times = 8)

for(i in c(1, 2, 5,6)) {
    kappa[i] <- round(eval(parse(text=paste0("fit_model_", i, "$kappa"))) , 3)}


d   <- seq(0, 30, length.out = 200)
cor <- matrix(-99, nrow = length(d), 8)
for (i in 1:8)
    {
    cov <- c(nugget[i], rep(0, times = (length(d) - 1) ) ) + partial_sill[i] *
            matern(d, phi = phi[i], kappa = kappa[i])
    cor[ , i] <- cov / cov[1]
    }

data.frame(
    d   = rep(d[-1], times = 8),
    cor = c(cor[-1, ]),
    model = rep(paste0("Model_", 1:8), each = (length(d)-1) )
) %>%

    ggplot(aes(x = d, y = cor)) +
        geom_line() +
        geom_point(data = data.frame(d_nugget = d[1], cor_nugget = cor[1]), 
                aes(x = d_nugget, y = cor_nugget), inherit.aes = FALSE) +                                    facet_wrap(~model, ncol= 4) +
        ggtitle("Figure 4: Estimated Correlation Function") + 
        theme(plot.title = element_text(hjust = 0.5)) +
        xlab("Distance") +
        ylab("Correlation") +
        ylim(c(0, max(cor)))
```
Figure 4 depicts the fitted correlation function for each of the 8 candidate models. Models in the first row have spatial covariates and models in the second row have intercept only. The figure shows that after adding the spatial covariates the correlation drops faster compared to the models without spatial covariates. And in Model 2 (the best model) the decay is the fastest.

# Problem 4

Fit a non-spatial regression (e.g., use the _lm_ function). Look at the regression coefficient estimates. Is there evidence that the inclusion of the spatial model is changing the covariate estimates? **Write a few sentences to interpret the results.** Make sure you write out the answers in **clear, English sentences. Communication matters!**

\textbf{Answer:}

```{r}
knitr::kable(
    t(data.frame(
        Model      = 1:8, 
        Covariates = rep(c("Spat_Coef", "Inpt_only"), each = 4),
        Covariance = rep(rep(c("Matern", "Exp"), each = 2), times = 2),
        Nugget     = nugget,
        beta_0     = beta_0,
        beta_1     = beta_1,
        beta_2     = beta_2,
        beta_3     = beta_3,
        part_sill     = partial_sill,
        phi     = phi,
        kappa     = kappa
    ) ), 
    caption = "Estimated values of the parameters for all models (Here zero represents 
    the corresponding parameter is fixed at zero in the corresponding model)"
) 

```


```{r}
dat %>%
    do(broom::tidy(lm(log_annual_total ~ LAT + LONG + ALT, .)))
```
Table 3 shows the estimated values of the parameters of all candidate models. Table 2 and the non-spatial regression outputs in the above clearly shows that inclusion of spatial model decrease all the regression coefficients like the intercept drops by about 6 units form around 29, the coefficients of latitude, longitude becomes around -0.10 and -0.37 from -0.06 and -0.31 after including spatial model. The coefficient of altitude remains similar in both cases.


# Problem 5

For the model with the smallest BIC, fit separate models for sites in the northern region vs. the southern region (North of 37.20 degrees latitude (inclusive) vs. south of 37.20 degrees latitude (exclusive)). Based on this fit, do you think the residual covariance is stationary? **Write a few sentences to interpret the results.** Make sure you write out the answers in **clear, English sentences. Communication matters!**

\textbf{Answer:}


```{r}

dat_north <- dat %>%
                subset(LAT >= 37.2)

dat_south <- dat %>%
                subset(LAT < 37.2)
    
fit_model_2_north <- likfit(
    data         = dat_north$log_annual_total,
    trend        = ~ dat_north$LAT + dat_north$LONG + dat_north$ALT,
    coords       = cbind(dat_north$LAT, dat_north$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat_north$log_annual_total), .5),
    message = FALSE
)

#summary(fit_model_2_north)

fit_model_2_south <- likfit(
    data         = dat_south$log_annual_total,
    trend        = ~ dat_south$LAT + dat_south$LONG + dat_south$ALT,
    coords       = cbind(dat_south$LAT, dat_south$LONG),
    fix.kappa = FALSE, kappa = 1,
    cov.model    = "matern",
    ini.cov.pars = c(var(dat_south$log_annual_total), .5),
    message = FALSE
)

nugget_new <- round(c(fit_model_2_north$nugget, fit_model_2_south$nugget), 4)
partial_sill_new <- round( c(fit_model_2_north$sigmasq, fit_model_2_south$sigmasq),3 )
phi_new <- round( c(fit_model_2_north$phi, fit_model_2_south$phi),3 )
kappa_new <- round(c(fit_model_2_north$kappa, fit_model_2_south$kappa), 3)

knitr::kable(
    data.frame(
        Model      = c("Model_2", "Model_2_North", "Model_2_South" ), 
        Nugget     = c(nugget[2], nugget_new),
        Partial_sill  = c(partial_sill[2],  partial_sill_new),
        Phi        = c(phi[2], phi_new),
        Kappa      = c(kappa[2], kappa_new)
    ), 
    caption = "Estimated values of the covariance parameters"
) 

d   <- seq(0, 5, length.out = 200)
cov <- matrix(-99, nrow = length(d), 2)
for (i in 1:2)
    {
    cov[ , i] <- c(nugget_new[i], rep(0, times = (length(d) - 1) ) ) + 
        partial_sill_new[i]*matern(d, phi = phi_new[i], kappa = kappa_new[i])
    }

cov = cbind( c(nugget[2], rep(0, times = (length(d) - 1) ) ) + 
                 partial_sill[2]*matern(d, phi = phi[2], kappa = kappa[2]), cov )

data.frame(
    d   = rep(d[-1], times = 3),
    cov = c(cov[-1, ]),
    Model = rep(c("Model_2", "Model_2_North", "Model_2_South" ), each = (length(d)-1) )
    ) %>%
    ggplot(aes(x = d, y = cov, col = Model)) +
        geom_line() +
        geom_point(data = data.frame(d_nugget = rep(d[1], 3), cov_nugget = cov[1, ], 
                          Model = c("Model_2", "Model_2_North", "Model_2_South" ) ), 
                aes(x = d_nugget, y = cov_nugget, col = Model), inherit.aes = FALSE) +                                    ggtitle("Figure 5: Estimated Covariance Function") + 
        ylim(c(0, max(cov))) +
        xlab("Distance") +
        ylab("Covariance") +
        theme(plot.title = element_text(hjust = 0.5))
```

Table 4 shows that the nugget and partial sill are similar in both north as well as in south part, and these values are also close the corresponding value in the full model. However, in the south part the phi and kappa takes slightly lower values compared to the corresponding estimation in the north. The covariance function plot in Figure 5 also depicts that the patterns are close in both part. Therefore, we can interpret that the residual covariance are almost stationary. 




